DRAFT: v8 \\ Vibration Characterisation of Strollers and\\ Cargo Bicycles for Transporting Infants\\ \vspace{2mm} \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\

### 2025-10-21 main.tex
Version 8, added ALT TEXT to each image, author contrib statement, and practitioner summary.

\title{
  DRAFT: v7 \\
  Vibration Characterisation of Strollers and\\
  Cargo Bicycles for Transporting Infants\\
  \vspace{2mm}
  \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\
}

@@

\color{black}
\begin{abstract}
  % NOTE: keep to 150 words max
  % Knowledge is lacking on road-induced vibrations for infants transported  in
  % strollers and cargo  bicycles.
  We evaluated vibrations with dummies representing infants aged 0, 3, and 9
  months lying or sitting in five strollers and two cargo bicycles with
  dedicated baby seats on six common road surfaces using the ISO standard for
  whole-body vibration.
  Strollers induced on average 0.4~\si{\mps\squared} on tarmac and up to
  5.0~\si{\mps\squared} on cobblestones at a mean walking speed of
  5.3~\si{\kph}.
  Cargo bicycles induced on average 0.6~\si{\mps\squared} on tarmac and up to
  10.7~\si{\mps\squared} at 25~\si{\kph} on paver bricks.
  The standard suggests the highest accelerations for strollers and cargo
  bicycles are extremely uncomfortable and continuous exposure should be limited
  to less than 10~\si{\minute}.
  % Effects of body size and posture were minor.
  Two vintage strollers significantly reduced vibrations compared to three
  modern strollers, indicating benefits of compliant suspensions. 
  We recommend designers systematically consider vibration, users avoid
  prolonged exposure to surfaces rougher than tarmac, and researchers to pursue
  scientifically founded test procedures and standards for infant vibration.
\end{abstract}

@@

\section{Materials and Methods}
%
\subsection{Test Equipment}
%
Vibration measurements were conducted using five different strollers and two
cargo bicycles with two baby seats, chosen for their popularity and/or
distinctive features. The strollers included three modern strollers:
\underline{Bugaboo} Fox 5 (Bugaboo, Amsterdam, The Netherlands),
\underline{Maxi-Cosi} Street Plus (Dorel Juvenile, Foxborough, USA), and
\underline{Stokke} BABYZEN YOYO 0+ (Stokke, Ålesund, Norway) and two vintage
strollers from unknown manufacturers ``\underline{Green Machine}'' cot-style and
``\underline{Old Rusty}'' seat-style. The cargo bicycles were a delta
configuration tricycle \underline{Keiler} (Keiler Bakfiets, The Netherlands) and
a two-wheel \underline{Urban Arrow} (Pon Bicycle Holding B.V., Amsterdam, The
Netherlands). Each was tested with both seats: Maxi-Cosi X Joolz
\underline{Pebble} Pro i-Size (Dorel Juvenile, Foxborough, USA) and
\underline{Melia} Baby Shell (Melia, Rotterdam, The Netherlands). The underlined
words for each product designate an abbreviated name used in figures and tables
in the paper and Figure~\ref{fig:equipment} shows the tested
products.~\footnote{Appendix~\ref{app:equipment} provides more detailed figures
and technical descriptions of all tested strollers and cargo bicycles.}
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/equipment.png}
  \caption{Tested products: cargo bicycles with baby seats, strollers, and
  dummies. Images are reproduced from manufacturer's websites except for the
  vintage strollers and dummies.}
  \label{fig:equipment}
\end{figure}

@@

\subsection{Measurement Equipment}
%
Linear accelerations and angular velocities were measured tri-axially using
Consensys Shimmer3 IMUs with a Consensys Base 6U.01 dock (Shimmer Sensing,
Dublin, Ireland). The IMUs were updated to firmware version LogAndStream v0.11.0
and managed via the software ConsensysBASIC v1.6.0-64bit on a Dell 7310 laptop
with Microsoft Windows 10. We 3D printed supports (material: PLA) for the
sensors to firmly fix the sensors on the tested bicycle/stroller (small white
and orange boxes visible in Figure~\ref{fig:sensors_UA}). Each vehicle was
equipped with five IMUs. The IMUs under the head and buttocks contacting the
dummy were placed according to the recommended practice in vibration testing
standard ISO-2631-1.~\footnote{Only two sensors were used in this study and
descriptions of the remaining three sensors and detailed drawings showing the
vehicles and the sensors' location are in Appendix~\ref{app:equipment}.} This
generates acceleration data representative of the mechanical load transferred
from the seat to the human body taking into account the compliance of the seat
foams and the inertia and compliance of the dummy. 
%
\begin{figure}
  \centering
  \subcaptionbox{IMU on Urban Arrow rear hub.}{
    \includegraphics[width=65mm, angle=-90]{fig/RW_UA.jpg}}
  \subcaptionbox{IMUs on the Pebble seat.}{
    \includegraphics[width=65mm, angle=-90]{fig/Maxicosi_sensors_UA.jpg}}
  \subcaptionbox{Stokke, dummy, and camera.}{
    \includegraphics[height=65mm]{fig/stroller-dummy-camera.jpg}}
  \caption{Example equipment set up. a) An IMU was placed on the wheel hub
  (cargo bicycles) or clamped to the wheel (strollers) for measuring travel
  speed. The travel (longitudinal) speed was derived according to the wheel
  radius and the IMU's angular speed about the wheel axis. In b) the IMUs were
  taped into the baby seat at the interface between the dummy's buttocks or head
  and the baby seat. In c) the camera is mounted to the stroller handlebar and
  IMUs are on the wheels.}
  \label{fig:sensors_UA}
\end{figure}

@@

\subsection{Experimental Protocol}
%
All tests were conducted in Delft, The Netherlands, on public roads near Delft
University of Technology~\footnote{Details of all test locations and surfaces
are shown in Appendix~\ref{app:location}.}. After mounting the sensors and
reaching the location of the experiment, we turned on the sensors to start the
experimental ``session'', where a session is a continuous data collection period
from a single vehicle that may include different road surfaces or speeds. Before
each trial, we pushed the vehicle back and forth on level ground to mark the
beginning of the session as an extra time-synchronization measure. For the cargo
bicycle, a single rider conducted all trials to be consistent across different
sessions (rider mass: 59~\si{\kg}; rider height: 1.70~\si{\m}). We tested
cargo bicycles on tarmac and paver bricks road surfaces at target speeds of
12~\si{\kph}, 20~\si{\kph}, and 25~\si{\kph}. The stroller tests were conducted
on tarmac, paver bricks, sidewalk pavers, cobblestones, and sidewalk slabs
(concrete blocks with gaps in between) road surfaces. Figure~\ref{fig:surfaces}
shows the road surfaces used for each vehicle type. In all tests, the speed was
manually controlled by the pusher or cyclist by observing a speedometer mounted
on the handlebar of the stroller or cargo bicycle. 
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/surfaces.png}
  \caption{Road surfaces tested: tarmac and paver bricks (cargo bicycles and
  strollers), sidewalk pavers, cobblestones, sidewalk slabs (strollers).}
  \label{fig:surfaces}
\end{figure}

@@

Each session was segmented into ``trials'' corresponding to a different road
surface or activity and each trial was divided into subsequent, back-to-back
``repetitions''. We divided into repetitions to capture the variation of road
surface features within a trial. The raw data consists of a single
comma-separated value (CSV) file per session per IMU, along with metadata for
the sessions and vehicles. The CSV file contains the time series data from each
IMU: linear acceleration along and angular speed about each of the body-fixed
orthogonal axes of the IMU alongside Epoch Unix timestamps. We segmented the
sessions into trials representing a motion state of the vehicle: either static
on level ground or being propelled over one of the surfaces of interest at a
constant speed. Figure~\ref{fig:session} gives an example of segmenting a
session in different trials. The data were processed per session as follows:
%
% NOTE : Putting the figure before the enumeration causes a slightly better
% layout.
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/session015.png}
  \caption{Vehicle speed versus time derived from the wheel hub IMU for an
  entire session (015) showing the trials in the shaded grey areas. Gold
  horizontal lines depict the mean speed during trials bounded by its standard
  deviation.}
  \label{fig:session}
\end{figure}
%
\begin{enumerate}
  \item Calculate the vehicle travel speed during the session from the angular
    rate of the rear wheel and the vehicle's wheel radius.
  \item Extract segments representing a single trial from each session time
    history, based on the manually labelled segment ``start'', and ``end'' times.
  \item Split the trials with durations longer than 40~\si{\second} into
    repetitions of \SIrange{20}{39}{\second}. Trials less than 20~\si{\second}
    are not split.
  \item Rotate the accelerometer data for each sensor from body-fixed sensor
    coordinates to body-fixed vehicle coordinates. This is achieved by rotating
    the coordinate axes about the sensor's body-fixed axis, which was manually
    aligned with the vehicle's pitch axis. We subtracted the mean measured
    acceleration due to gravity (standard gravity) giving linear acceleration of
    each sensor projected into the vehicle's SAE body-fixed
    axes~\cite{SocietyofAutomotiveEngineers2008}, named: longitudinal $x$,
    lateral $y$, and vertical $z$.
  \item Extract each motion trial segment and select the vehicle body-fixed
    longitudinal, lateral, and vertical component of the seat pan accelerometer.
\end{enumerate}
%
This resulted in acceleration versus time recordings of 154 total repetitions.
The repetitions have durations in the range of \SIrange{20}{40}{\second} (mean:
25~\si{\second}), see Table~\ref{tab:num-trials}.

@@

\subsection{Data Analysis}
%
Figure~\ref{fig:vert-acc-example} shows an example time history of the vertical
(\(z\)) acceleration of the seat pan during a single repetition. To perform the
ISO~2631-1 recommended health and comfort analysis, we first downsampled the
time history from the hardware-set variable sampling frequency of approximately
910.22~\si{\hertz} to a constant sample rate of 400~\si{\hertz} using linear
interpolation, giving sufficient samples for the bandwidth of interest based on
the Nyquist frequency. We set any acceleration values outside of the sensor
manufacturer's reported operating range of \(\pm\)16~g to that maximum or
minimum, respectively, given that values outside the range may be unreliable.
Values that exceeded the range only present rarely in nine of the cargo bicycle
paver brick repetitions. Following the ISO~2631-1 recommendation, we low-pass
filtered the signal at \(1.5\times80~\si{\hertz}=120~\si{\hertz}\) using a
zero-lag 2\textsuperscript{nd} order Butterworth filter, given that the standard
only applies to frequencies up to 80~\si{\hertz}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/session001-t2-aula-stroller-maxicosi-cot-0-SeatBotacc_ver-rep0.png}
  \caption{Raw seat pan vertical acceleration versus time from session 001:
  Maxi-Cosi stroller over the Sidewalk Slabs. Black horizontal lines indicate
  the \(\pm\)root mean square (RMS) about the mean. The vibration dose value
  (VDV) for the first 10~\si{\second} of the printed duration is shown in the
  bottom left.}
  \label{fig:vert-acc-example}
\end{figure}

ISO~2631-1 provides weighting filters that highlight frequencies that adults are
most sensitive to. To apply them, we calculated the amplitude spectra of the
acceleration time histories of each trial using the Fast Fourier Transform
(FFT). Figure~\ref{fig:freq-spectrum} gives an example of a raw amplitude
spectrum along with smoothed versions of the raw and ISO~2631-1 weighted
signals.  In almost all repetitions, there is a single dominant peak frequency
in the ISO weighted and smoothed spectra. A handful of trials had two or more
peaks at adjacent frequencies of approximately the same amplitude. We selected
the maximum amplitude peak in those cases. Before ISO weighting, the area under
the spectrum curve was calculated and the frequency below which 80\% of the
amplitude content falls marked as an indicator of bandwidth.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/session004-t0-pave-stroller-maxicosi-cot-0-SeatBotacc_ver-rep0.png}
  \caption{Seat pan vertical acceleration amplitude spectrum from session 004:
  Maxi-Cosi stroller over cobblestones. The gray curve shows the result of
  the FFT, the black line is a smoothened version of the FFT (zero-lag Butterworth
  low pass), and the blue line is the smoothed ISO weighted FFT. The blue dashed
  vertical line indicates the frequency at the maximum amplitude of the smoothed
  curve. The black dashed-dotted vertical line indicates the bandwidth threshold
  for 80\% of the area under the black curve.}
  \label{fig:freq-spectrum}
\end{figure}

@@

\subsection{Effect of Speed and model with Cargo Bicycles}
%
Figure~\ref{fig:bicycle-type-compare} compares the two cargo bicycle models
which were both fitted with the same set of two baby seats (Melia and Pebble).
Keiler sees higher accelerations compared to the two-wheeled Urban Arrow, with a
pronounced increase on tarmac and a modest increase on paver bricks. Both
vehicles show increasing accelerations with speed.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-bicycle-type-compare.png}
  \caption{Seat pan ISO~2631-1 weighted vertical RMS acceleration versus speed
  grouped by road surface and cargo bicycle model. Slanted lines indicate a
  linear regression, vertical lines are the standard deviation at those speeds,
  and shaded regions show the 95\% confidence intervals for the regression.}
  \label{fig:bicycle-type-compare}
\end{figure}

\subsection{Effect of Stroller Model}
%
Figure~\ref{fig:stroller-type-compare} shows the vertical RMS accelerations for
each road surface for each of the five strollers, lumping seat configurations
and dummy sizes. The Maxi-Cosi and Stokke strollers have similar mean values.
The Bugaboo has a slightly lower mean for cobblestones, paver bricks, and
sidewalk pavers. The Green Machine performs better than the modern strollers on
paver bricks, but similarly otherwise. The Old Rusty performs better than the
modern strollers on all surfaces except tarmac. All strollers seem to experience
similar accelerations on tarmac. All road surfaces compared to tarmac at least
double the RMS acceleration.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-stroller-type-compare.png}
  \caption{Seat pan ISO 2631-1 weighted vertical RMS acceleration per road
  surface for each stroller. Vertical lines indicate the standard deviation for
  categories that have more than one repetition.} 
  \label{fig:stroller-type-compare}
\end{figure}

\subsection{Dominant Frequency and Bandwidth}
%
Figure~\ref{fig:spectra-compare} shows frequency spectra averaged over body size
and posture for all strollers on sidewalk pavers (left) and cargo bicycles on
paver bricks. These surfaces are highly relevant as they are common in the
Netherlands. Apparently, the two vintage strollers show a lower peak frequency,
which is 4.1~\si{\hertz} for Green Machine and 5.6~\si{\hertz} for Old Rusty
whereas modern systems peak around \SIrange{7}{9}{\hertz}. This can be explained
by the more compliant suspension of the vintage strollers. The Old Rusty shows
the lowest acceleration peak, and the lowest RMS weighted acceleration, but
above 30 Hz it shows the highest power of all strollers. The Keiler sees much
higher peak values as compared to the Urban Arrow. In both cargo bicycles, the
peak frequencies range from \SIrange{6}{10}{\hertz} and hardly depend on vehicle
and speed. Figure~\ref{fig:peak-freq-dist} shows the distributions of the
dominant (peak) frequency across road surface types for each of the target
speeds. Peak frequencies range from about \SIrange{4}{11}{\hertz} across all
trials. For strollers (5~\si{\kph}), the median frequency increases from
sidewalk slabs to cobblestones and sidewalk pavers and then to tarmac and paver
bricks. For cargo bicycles, the difference in peak frequency between the two
road surfaces is not as apparent or consistent.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-spectra-compare.png}
  \caption{Mean vertical seat pan amplitude spectra of each vehicle for
  strollers on sidewalk pavers (left) and cargo bicycles on paver bricks
  (right).} 
  \label{fig:spectra-compare}
\end{figure}
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-peak-freq-dist.png}
  \caption{Seat pan vertical peak frequency distributions comparisons among road surfaces for each
  target speed group for non-shock repetitions. The boxes bound the quartiles
  and indicate the median. The whiskers indicate the 95\textsuperscript{th}
  percentile and circles are outliers.}
  \label{fig:peak-freq-dist}
\end{figure}

Figure~\ref{fig:bandwidth-dist} gives a general indication of the bandwidth (80\% of
the amplitude spectrum content) for each of the target speed groups. On average,
the bandwidth is 56 Hz for modern strollers, 46 Hz for Green Machine, 105 Hz for
Old Rusty, and 44 Hz for cargo bicycles.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-bandwidth-dist.png}
  \caption{Bandwidth (based on 80\% of the area under the unfiltered seat pan
  vertical amplitude spectrum) for each target speed group for non-shock
  repetitions. The boxes bound the quartiles and indicate the median. The
  whiskers indicate the 95\textsuperscript{th} percentile and circles are
  outliers.}
  \label{fig:bandwidth-dist}
\end{figure}

>>>>

\title{
  DRAFT: v8 \\
  Vibration Characterisation of Strollers and\\
  Cargo Bicycles for Transporting Infants\\
  \vspace{2mm}
  \large{Including Recommendations for Users, Designers, Manufacturers, and Researchers} \\
}

@@

\color{black}
\begin{abstract}
  % NOTE: keep to 150 words max
  % Knowledge is lacking on road-induced vibrations for infants transported  in
  % strollers and cargo  bicycles.
  We evaluated vibrations with dummies representing infants aged 0, 3, and 9
  months lying or sitting in five strollers and two cargo bicycles with
  dedicated baby seats on six common road surfaces using the ISO standard for
  whole-body vibration.
  Strollers induced on average 0.4~\si{\mps\squared} on tarmac and up to
  5.0~\si{\mps\squared} on cobblestones at a mean walking speed of
  5.3~\si{\kph}.
  Cargo bicycles induced on average 0.6~\si{\mps\squared} on tarmac and up to
  10.7~\si{\mps\squared} at 25~\si{\kph} on paver bricks.
  The standard suggests the highest accelerations for strollers and cargo
  bicycles are extremely uncomfortable and continuous exposure should be limited
  to less than 10~\si{\minute}.
  % Effects of body size and posture were minor.
  Two vintage strollers significantly reduced vibrations compared to three
  modern strollers, indicating benefits of compliant suspensions. 
  We recommend that designers systematically consider vibration, users avoid
  prolonged exposure to surfaces rougher than tarmac, and researchers pursue
  scientifically founded test procedures and standards for infant vibration.
\end{abstract}

\section*{Practitioner Summary}
%
Strollers pushed over cobblestones and cargo bicycles travelling over paver
bricks at typical speeds induce infant seat pan accelerations up to 5 and
11~\si{\mps\squared}, respectively. The adult whole-body vibration standard
suggests these accelerations are extremely uncomfortable and continuous exposure
should be limited to less than 10~\si{\minute}.

@@

\section{Materials and Methods}
%
\subsection{Test Equipment}
%
Vibration measurements were conducted using five different strollers and two
cargo bicycles with two baby seats, chosen for their popularity and/or
distinctive features. The strollers included three modern strollers:
\underline{Bugaboo} Fox 5 (Bugaboo, Amsterdam, The Netherlands),
\underline{Maxi-Cosi} Street Plus (Dorel Juvenile, Foxborough, USA), and
\underline{Stokke} BABYZEN YOYO 0+ (Stokke, Ålesund, Norway) and two vintage
strollers from unknown manufacturers ``\underline{Green Machine}'' cot-style and
``\underline{Old Rusty}'' seat-style. The cargo bicycles were a delta
configuration tricycle \underline{Keiler} (Keiler Bakfiets, The Netherlands) and
a two-wheel \underline{Urban Arrow} (Pon Bicycle Holding B.V., Amsterdam, The
Netherlands). Each was tested with both seats: Maxi-Cosi X Joolz
\underline{Pebble} Pro i-Size (Dorel Juvenile, Foxborough, USA) and
\underline{Melia} Baby Shell (Melia, Rotterdam, The Netherlands). The underlined
words for each product designate an abbreviated name used in figures and tables
in the paper and Figure~\ref{fig:equipment} shows the tested
products.~\footnote{Appendix~\ref{app:equipment} provides more detailed figures
and technical descriptions of all tested strollers and cargo bicycles.}
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/equipment.png}
  \caption{Tested products: cargo bicycles with baby seats, strollers, and
  dummies. Images are reproduced from manufacturer's websites except for the
  vintage strollers and dummies. ALT TEXT: Photos of cargo bicycles, infant
  seats, strollers, and infant dummies.}
  \label{fig:equipment}
\end{figure}

@@

\subsection{Measurement Equipment}
%
Linear accelerations and angular velocities were measured tri-axially using
Consensys Shimmer3 IMUs with a Consensys Base 6U.01 dock (Shimmer Sensing,
Dublin, Ireland). The IMUs were updated to firmware version LogAndStream v0.11.0
and managed via the software ConsensysBASIC v1.6.0-64bit on a Dell 7310 laptop
with Microsoft Windows 10. We 3D printed supports (material: PLA) for the
sensors to firmly fix the sensors on the tested bicycle/stroller (small white
and orange boxes visible in Figure~\ref{fig:sensors_UA}). Each vehicle was
equipped with five IMUs. The IMUs under the head and buttocks contacting the
dummy were placed according to the recommended practice in vibration testing
standard ISO-2631-1.~\footnote{Only two sensors were used in this study and
descriptions of the remaining three sensors and detailed drawings showing the
vehicles and the sensors' location are in Appendix~\ref{app:equipment}.} This
generates acceleration data representative of the mechanical load transferred
from the seat to the human body taking into account the compliance of the seat
foams and the inertia and compliance of the dummy. 
%
\begin{figure}
  \centering
  \subcaptionbox{IMU on Urban Arrow rear hub.}{
    \includegraphics[width=65mm, angle=-90]{fig/RW_UA.jpg}}
  \subcaptionbox{IMUs on the Pebble seat.}{
    \includegraphics[width=65mm, angle=-90]{fig/Maxicosi_sensors_UA.jpg}}
  \subcaptionbox{Stokke, dummy, and camera.}{
    \includegraphics[height=65mm]{fig/stroller-dummy-camera.jpg}}
  \caption{Example equipment set up. a) An IMU was placed on the wheel hub
  (cargo bicycles) or clamped to the wheel (strollers) for measuring travel
  speed. The travel (longitudinal) speed was derived according to the wheel
  radius and the IMU's angular speed about the wheel axis. In b) the IMUs were
  taped into the baby seat at the interface between the dummy's buttocks or head
  and the baby seat. In c) the camera is mounted to the stroller handlebar and
  IMUs are on the wheels. ALT TEXT: Photos of IMU and camera locations on the
  vehicles.}
  \label{fig:sensors_UA}
\end{figure}

@@

\subsection{Experimental Protocol}
%
All tests were conducted in Delft, The Netherlands, on public roads near Delft
University of Technology~\footnote{Details of all test locations and surfaces
are shown in Appendix~\ref{app:location}.}. After mounting the sensors and
reaching the location of the experiment, we turned on the sensors to start the
experimental ``session'', where a session is a continuous data collection period
from a single vehicle that may include different road surfaces or speeds. Before
each trial, we pushed the vehicle back and forth on level ground to mark the
beginning of the session as an extra time-synchronization measure. For the cargo
bicycle, a single rider conducted all trials to be consistent across different
sessions (rider mass: 59~\si{\kg}; rider height: 1.70~\si{\m}). We tested
cargo bicycles on tarmac and paver bricks road surfaces at target speeds of
12~\si{\kph}, 20~\si{\kph}, and 25~\si{\kph}. The stroller tests were conducted
on tarmac, paver bricks, sidewalk pavers, cobblestones, and sidewalk slabs
(concrete blocks with gaps in between) road surfaces. Figure~\ref{fig:surfaces}
shows the road surfaces used for each vehicle type. In all tests, the speed was
manually controlled by the pusher or cyclist by observing a speedometer mounted
on the handlebar of the stroller or cargo bicycle. 
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/surfaces.png}
  \caption{Road surfaces tested: tarmac and paver bricks (cargo bicycles and
  strollers), sidewalk pavers, cobblestones, sidewalk slabs (strollers). ALT
  TEXT: Photos of road surfaces.}
  \label{fig:surfaces}
\end{figure}

@@

Each session was segmented into ``trials'' corresponding to a different road
surface or activity and each trial was divided into subsequent, back-to-back
``repetitions''. We divided into repetitions to capture the variation of road
surface features within a trial. The raw data consists of a single
comma-separated value (CSV) file per session per IMU, along with metadata for
the sessions and vehicles. The CSV file contains the time series data from each
IMU: linear acceleration along and angular speed about each of the body-fixed
orthogonal axes of the IMU alongside Epoch Unix timestamps. We segmented the
sessions into trials representing a motion state of the vehicle: either static
on level ground or being propelled over one of the surfaces of interest at a
constant speed. Figure~\ref{fig:session} gives an example of segmenting a
session in different trials. The data were processed per session as follows:
%
% NOTE : Putting the figure before the enumeration causes a slightly better
% layout.
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/session015.png}
  \caption{Vehicle speed versus time derived from the wheel hub IMU for an
  entire session (015) showing the trials in the shaded grey areas. Gold
  horizontal lines depict the mean speed during trials bounded by its standard
  deviation. ALT TEXT: Plot of vehicle speed during a session.}
  \label{fig:session}
\end{figure}
%
\begin{enumerate}
  \item Calculate the vehicle travel speed during the session from the angular
    rate of the rear wheel and the vehicle's wheel radius.
  \item Extract segments representing a single trial from each session time
    history, based on the manually labelled segment ``start'', and ``end'' times.
  \item Split the trials with durations longer than 40~\si{\second} into
    repetitions of \SIrange{20}{39}{\second}. Trials less than 20~\si{\second}
    are not split.
  \item Rotate the accelerometer data for each sensor from body-fixed sensor
    coordinates to body-fixed vehicle coordinates. This is achieved by rotating
    the coordinate axes about the sensor's body-fixed axis, which was manually
    aligned with the vehicle's pitch axis. We subtracted the mean measured
    acceleration due to gravity (standard gravity) giving linear acceleration of
    each sensor projected into the vehicle's SAE body-fixed
    axes~\cite{SocietyofAutomotiveEngineers2008}, named: longitudinal $x$,
    lateral $y$, and vertical $z$.
  \item Extract each motion trial segment and select the vehicle body-fixed
    longitudinal, lateral, and vertical component of the seat pan accelerometer.
\end{enumerate}
%
This resulted in acceleration versus time recordings of 154 total repetitions.
The repetitions have durations in the range of \SIrange{20}{40}{\second} (mean:
25~\si{\second}), see Table~\ref{tab:num-trials}.

@@

\subsection{Data Analysis}
%
Figure~\ref{fig:vert-acc-example} shows an example time history of the vertical
(\(z\)) acceleration of the seat pan during a single repetition. To perform the
ISO~2631-1 recommended health and comfort analysis, we first downsampled the
time history from the hardware-set variable sampling frequency of approximately
910.22~\si{\hertz} to a constant sample rate of 400~\si{\hertz} using linear
interpolation, giving sufficient samples for the bandwidth of interest based on
the Nyquist frequency. We set any acceleration values outside of the sensor
manufacturer's reported operating range of \(\pm\)16~g to that maximum or
minimum, respectively, given that values outside the range may be unreliable.
Values that exceeded the range only present rarely in nine of the cargo bicycle
paver brick repetitions. Following the ISO~2631-1 recommendation, we low-pass
filtered the signal at \(1.5\times80~\si{\hertz}=120~\si{\hertz}\) using a
zero-lag 2\textsuperscript{nd} order Butterworth filter, given that the standard
only applies to frequencies up to 80~\si{\hertz}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/session001-t2-aula-stroller-maxicosi-cot-0-SeatBotacc_ver-rep0.png}
  \caption{Raw seat pan vertical acceleration versus time from session 001:
  Maxi-Cosi stroller over the Sidewalk Slabs. Black horizontal lines indicate
  the \(\pm\)root mean square (RMS) about the mean. The vibration dose value
  (VDV) for the first 10~\si{\second} of the printed duration is shown in the
  bottom left. ALT TEXT: Plot of acceleration versus time during an example
  repetition.}
  \label{fig:vert-acc-example}
\end{figure}

ISO~2631-1 provides weighting filters that highlight frequencies that adults are
most sensitive to. To apply them, we calculated the amplitude spectra of the
acceleration time histories of each trial using the Fast Fourier Transform
(FFT). Figure~\ref{fig:freq-spectrum} gives an example of a raw amplitude
spectrum along with smoothed versions of the raw and ISO~2631-1 weighted
signals.  In almost all repetitions, there is a single dominant peak frequency
in the ISO weighted and smoothed spectra. A handful of trials had two or more
peaks at adjacent frequencies of approximately the same amplitude. We selected
the maximum amplitude peak in those cases. Before ISO weighting, the area under
the spectrum curve was calculated and the frequency below which 80\% of the
amplitude content falls marked as an indicator of bandwidth.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/session004-t0-pave-stroller-maxicosi-cot-0-SeatBotacc_ver-rep0.png}
  \caption{Seat pan vertical acceleration amplitude spectrum from session 004:
  Maxi-Cosi stroller over cobblestones. The gray curve shows the result of
  the FFT, the black line is a smoothened version of the FFT (zero-lag Butterworth
  low pass), and the blue line is the smoothed ISO weighted FFT. The blue dashed
  vertical line indicates the frequency at the maximum amplitude of the smoothed
  curve. The black dashed-dotted vertical line indicates the bandwidth threshold
  for 80\% of the area under the black curve. ALT TEXT: Plot of amplitude versus
  frequency for an example repetition.}
  \label{fig:freq-spectrum}
\end{figure}

@@

\subsection{Effect of Speed and model with Cargo Bicycles}
%
Figure~\ref{fig:bicycle-type-compare} compares the two cargo bicycle models
which were both fitted with the same set of two baby seats (Melia and Pebble).
Keiler sees higher accelerations compared to the two-wheeled Urban Arrow, with a
pronounced increase on tarmac and a modest increase on paver bricks. Both
vehicles show increasing accelerations with speed.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-bicycle-type-compare.png}
  \caption{Seat pan ISO~2631-1 weighted vertical RMS acceleration versus speed
  grouped by road surface and cargo bicycle model. Slanted lines indicate a
  linear regression, vertical lines are the standard deviation at those speeds,
  and shaded regions show the 95\% confidence intervals for the regression. ALT
  TEXT: Plots of cargo bicycle acceleration versus speed for tarmac and paver
  bricks.}
  \label{fig:bicycle-type-compare}
\end{figure}

\subsection{Effect of Stroller Model}
%
Figure~\ref{fig:stroller-type-compare} shows the vertical RMS accelerations for
each road surface for each of the five strollers, lumping seat configurations
and dummy sizes. The Maxi-Cosi and Stokke strollers have similar mean values.
The Bugaboo has a slightly lower mean for cobblestones, paver bricks, and
sidewalk pavers. The Green Machine performs better than the modern strollers on
paver bricks, but similarly otherwise. The Old Rusty performs better than the
modern strollers on all surfaces except tarmac. All strollers seem to experience
similar accelerations on tarmac. All road surfaces compared to tarmac at least
double the RMS acceleration.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-stroller-type-compare.png}
  \caption{Seat pan ISO 2631-1 weighted vertical RMS acceleration per road
  surface for each stroller. Vertical lines indicate the standard deviation for
  categories that have more than one repetition. ALT TEXT: Plot of stroller
  accelerations for each road surface.} 
  \label{fig:stroller-type-compare}
\end{figure}

\subsection{Dominant Frequency and Bandwidth}
%
Figure~\ref{fig:spectra-compare} shows frequency spectra averaged over body size
and posture for all strollers on sidewalk pavers (left) and cargo bicycles on
paver bricks. These surfaces are highly relevant as they are common in the
Netherlands. Apparently, the two vintage strollers show a lower peak frequency,
which is 4.1~\si{\hertz} for Green Machine and 5.6~\si{\hertz} for Old Rusty
whereas modern systems peak around \SIrange{7}{9}{\hertz}. This can be explained
by the more compliant suspension of the vintage strollers. The Old Rusty shows
the lowest acceleration peak, and the lowest RMS weighted acceleration, but
above 30 Hz it shows the highest power of all strollers. The Keiler sees much
higher peak values as compared to the Urban Arrow. In both cargo bicycles, the
peak frequencies range from \SIrange{6}{10}{\hertz} and hardly depend on vehicle
and speed. Figure~\ref{fig:peak-freq-dist} shows the distributions of the
dominant (peak) frequency across road surface types for each of the target
speeds. Peak frequencies range from about \SIrange{4}{11}{\hertz} across all
trials. For strollers (5~\si{\kph}), the median frequency increases from
sidewalk slabs to cobblestones and sidewalk pavers and then to tarmac and paver
bricks. For cargo bicycles, the difference in peak frequency between the two
road surfaces is not as apparent or consistent.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-spectra-compare.png}
  \caption{Mean vertical seat pan amplitude spectra of each vehicle for
  strollers on sidewalk pavers (left) and cargo bicycles on paver bricks
  (right). ALT TEXT: Plot of amplitude versus frequency for each stroller and
  cargo bicycle.} 
  \label{fig:spectra-compare}
\end{figure}
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-peak-freq-dist.png}
  \caption{Seat pan vertical peak frequency distributions comparisons among road surfaces for each
  target speed group for non-shock repetitions. The boxes bound the quartiles
  and indicate the median. The whiskers indicate the 95\textsuperscript{th}
  percentile and circles are outliers. ALT TEXT: Box plots of peak frequencies
  for strollers and bicycles on each road surface.}
  \label{fig:peak-freq-dist}
\end{figure}

Figure~\ref{fig:bandwidth-dist} gives a general indication of the bandwidth (80\% of
the amplitude spectrum content) for each of the target speed groups. On average,
the bandwidth is 56 Hz for modern strollers, 46 Hz for Green Machine, 105 Hz for
Old Rusty, and 44 Hz for cargo bicycles.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{fig/SeatBotacc_ver-bandwidth-dist.png}
  \caption{Bandwidth (based on 80\% of the area under the unfiltered seat pan
  vertical amplitude spectrum) for each target speed group for non-shock
  repetitions. The boxes bound the quartiles and indicate the median. The
  whiskers indicate the 95\textsuperscript{th} percentile and circles are
  outliers. ALT TEXT: Box plots of bandwidth for strollers and bicycles on each
  road surface.}
  \label{fig:bandwidth-dist}
\end{figure}
